The Code of Ethics

### 2016-04-13 pde_two/threshold.tex
update paper

\documentclass{hw}
\usepackage{amsmath,amssymb}

@@

\subsection{Solution to the Steady State Problem}
We are attempting to solve the ordinary differential equation
\[
u''=-u^{3}.
\]
Multiplying both sides of the equation gives us
\begin{align*}
u''u'&=-u'u^{3}\\
\int u''u'&=\int -u'u^{3}\\
{1\over2}(u')^{2}&= -{1\over4}u^{4}+c\\
2(u')^{2}&=-u^{4}+c_{1}.
\end{align*}
This first order equation does not lend itself to separation of variables as readily as we would like
(although it is possible if we want to deal with imaginary numbers). Instead, we will use a technique
called the \textit{shooting method} in order to determine a solution. The shooting method provides
us with a way to relax boundary conditions into initial conditions. Recall that $(5)$ provides us
with boundary conditions that arose from the steady state of $(1)$. We will leave the first boundary
condition alone and change the second boundary condition so that $2u'(0)^{2}=a^{4}$. This will allow
us to factor the resulting term:
\begin{align*}
u'(x)^2&={1\over2}(a^{4}-u^{4})\\
{\du\over\dx}&= \sqrt{{1\over2}(a^{2}+u^{2})(a^{2}-u^{2})}\\
{\dx\over\du}&= {1\over\sqrt{{1\over2}(a^{2}+u^{2})(a^{2}-u^{2})}}.
\end{align*}
Integrating both sides and substituting $s$ for $u$ gives us
\[x=\int_{0}^{u}{1\over\sqrt{{1\over2}(a^{2}+s^{2})(a^{2}-s^{2})}}\mathop{ds}.\]
We can now make another substitution for $s$ in terms of $t$. Letting
\[
s=\sqrt{{a^2t^2\over2(1-t^2/2)}}
\]
yields
\[
a^2+s^2={2a^2\over2(1-t^2/2)}\qquad\qquad\text{and}\qquad\qquad
a^2-s^2={a^2(1-t^2)\over(1-t^2/2)}.
\]
We also have
\[
\mathop{ds}={a(t-1)(t-2)\over2\sqrt{2}(1-t^2/2)^{3/2}}\dt,
\]
which gives us the new integral
\[
ax=\int_{0}^{\sqrt{\frac{\displaystyle2u}{\displaystyle a^2+u^2}}}{1\over\sqrt{(1-t^2)(1-t^2)}}.
\]
We may notice that this integral is one of the ways to define $sn(x,k)$. Rewriting it in terms
of our Jacobi elliptic function gives us
\[
x=\int_{0}^{sn(x,k)}{1\over\sqrt{(1-t^2)(1-t^2/2)}}.
\]

>>>>

\documentclass{hw}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

@@

\subsection{Solution to the Steady State Problem}
We are attempting to solve the ordinary differential equation
\[
u''=-u^{3}.
\]
Multiplying both sides of the equation gives us
\begin{align*}
u''u'&=-u'u^{3}\\
\int u''u'&=\int -u'u^{3}\\
{1\over2}(u')^{2}&= -{1\over4}u^{4}+c\\
2(u')^{2}&=-u^{4}+c_{1}.
\end{align*}
This first order equation does not lend itself to separation of variables as readily as we would like
(although it is possible if we want to deal with imaginary numbers). Instead, we will use a technique
called the \textit{shooting method} in order to determine a solution. The shooting method provides
us with a way to relax boundary conditions into initial conditions. Recall that $(5)$ provides us
with boundary conditions that arose from the steady state of $(1)$. We will leave the first boundary
condition alone and change the second boundary condition so that $2u'(0)^{2}=a^{4}$. This will allow
us to factor the resulting term:
\begin{align*}
u'(x)^2&={1\over2}(a^{4}-u^{4})\\
{\du\over\dx}&= \sqrt{{1\over2}(a^{2}+u^{2})(a^{2}-u^{2})}\\
{\dx\over\du}&= {1\over\sqrt{{1\over2}(a^{2}+u^{2})(a^{2}-u^{2})}}.
\end{align*}
Integrating both sides and substituting $s$ for $u$ gives us
\[x=\int_{0}^{u}{1\over\sqrt{{1\over2}(a^{2}+s^{2})(a^{2}-s^{2})}}\mathop{ds}.\]
We can now make another substitution for $s$ in terms of $t$. Letting
\[
s=\sqrt{{a^2t^2\over2(1-t^2/2)}}
\]
yields
\[
a^2+s^2={2a^2\over2(1-t^2/2)}\qquad\qquad\text{and}\qquad\qquad
a^2-s^2={a^2(1-t^2)\over(1-t^2/2)}.
\]
We also have
\[
\mathop{ds}={a(t-1)(t-2)\over2\sqrt{2}(1-t^2/2)^{3/2}}\dt,
\]
which gives us the new integral
\[
ax=\int_{0}^{\sqrt{\frac{\displaystyle2u}{\displaystyle a^2+u^2}}}{1\over\sqrt{(1-t^2)(1-t^2)}}.
\]
We may notice that this integral is one of the ways to define $sn(x,k)$. Rewriting it in terms
of our Jacobi elliptic function gives us
\[
x=\int_{0}^{sn(x,k)}{1\over\sqrt{(1-t^2)(1-t^2/2)}}.
\]
%% TODO -- Expand on this information
Finally, we have a solution $u(x)$
\[
u(x)={a\ sn(ax,\sqrt{1/2})\over\sqrt{2-sn^{2}(ax,\sqrt{1/2})}}.
\]
Recall that the first positive zero of $sn(x,k)$ is given by
\[
K=\int_{0}^{1}{1\over\sqrt{(1-t^2)(1-{t^2\over2})}}\dt.
\]
Then $u(x)=u_{s}(x)$ provided that $a=2K/\pi$. We can see the results via these Mathematica plots:
\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.5]{u_of_1_x}
\end{subfigure}
\hfill
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.5]{u_of_2_x}
\end{subfigure}
\caption{The graphs of $u_{s}(x)$ for $a=1$ and $a=2$}
\end{figure}

\noindent It appears that, for $a=1$, $u_{s}(x)$ behaves similar to $\sin{x}$.

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.5]{u_of_1_x}
\end{subfigure}
\hfill
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.5]{sinx}
\end{subfigure}
\caption{The graphs of $u_{s}(x)$ for $a=1$ and $\sin{x}$}
\end{figure}

\subsection{Average of the Steady State Solution}

If we can find the average temperature of the steady state, then we know that our function $u(x)$
has a global solution (otherwise there would be points in the range of $u$ for which $u$ blows up).
We will want to show that
\[
{1\over\pi}\int_{0}^{\pi}u(x)\dx={\sqrt{2}\over2}.
\]
To begin, let
\[
y=\int_{0}^{x}u(t)\dt
\]
and $v(y)=u^{2}(x)$ where $x=x(y)=y^{-1}(x)$.

### 2016-04-13 writing_in_ethics/paper_four.tex
update paper

The ACM code of ethics provides those in technical industries with guidelines to follow when making tough ethical decisions. One such ethical decision arose in January of 1986 when NASA's Challenger space shuttle exploded above Cape Canaveral due to the neglect of suggestions from various engineers who were part of the team. One engineer, Roger Boisjoly, testified against NASA and the the engineering company Morton Thiokol during the trial, despite the recommendations given Morton Thiokol's lawyers. Although Boisjoly broke some of the guidelines of the ACM code of Ethics, his actions were not unethical.

@@

When Boisjoly testified in court, he told the jury about the faulty O-ring inside the rocket. This action can be seen as ethical under the ACM Code. If Boisjoly had omitted the details and obeyed the orders given to him by Morton Thiokol's lawyers, he would have been violating the Code in several ways. One aspect of the Code of Ethics is that one should provide appropriate professional review. Boisjoly had already done this numerous times before the Challenger shuttle was destroyed. During his testimony, he presented the same information that was already known by the managers at NASA and Morton Thiokol. In addition, if Boisjoly had not told the panel what he knew, then he would have been omitting the truth. The Code of Ethics states that a professional must be honest about circumstances that may cause a conflict of interest. Omission of the truth, especially when dealing with loss of human life, would have certainly been a conflict of interest. One can argue that by not following the advice of Morton Thiokol's lawyers was in violation of the Code of Ethics, since he did not keep engineering work confidential. The Code, however, states that the obligation to keep work confidential can be changed when the law or other sections of the Code require that the information be divulged.

>>>>

The ACM code of ethics provides those in technical industries with guidelines to follow when making tough ethical decisions. One such ethical decision arose in January of 1986 when NASA's Challenger space shuttle exploded above Cape Canaveral due to the neglect of suggestions from various engineers who were part of the team. One engineer, Roger Boisjoly, testified against NASA and the the engineering company Morton Thiokol during the trial, despite the recommendations given Morton Thiokol's lawyers. Although Boisjoly broke some of the guidelines of the ACM code of Ethics, his actions were not ultimately unethical since he took steps to prevent the launch and remained honest throughout the resulting court trials.

@@

When Boisjoly testified in court, he told the jury about the faulty O-ring inside the rocket. This action can be seen as ethical under the ACM Code. If Boisjoly had omitted the details and obeyed the orders given to him by Morton Thiokol's lawyers, he would have been violating the Code in several ways. One aspect of the Code of Ethics is that one should provide appropriate professional review. Boisjoly had already done this numerous times before the Challenger shuttle was destroyed. During his testimony, he presented the same information that was already known by the managers at NASA and Morton Thiokol. In addition, if Boisjoly had not told the panel what he knew, then he would have been omitting the truth. The Code of Ethics states that a professional must be honest about circumstances that may cause a conflict of interest. Omission of the truth, especially when dealing with loss of human life, would have certainly been a conflict of interest. One can argue that by not following the advice of Morton Thiokol's lawyers was a violation of the Code of Ethics, since he did not keep engineering work confidential. The Code, however, states that the obligation to keep work confidential can be changed when the law or other sections of the Code require that the information be divulged.
